---
title: "ECON 144 Proj 2"
author: "Leonard Zhu"
date: "2024-11-14"
output: pdf_document
---

```{r setup, include=FALSE}
library(readxl)
library(forecast)
library(tseries)
library(vars)
library(ggplot2)
```

# **1. Modeling and Forecasting Trend, Seasonality, and Cycles**


In this section, I load the data, ensure the date is formatted correctly, and merge the datasets by date to align both NVIDIA and S&P 500 data series.

```{r}
# Load NVIDIA data
nvidia_data <- read_excel("~/Downloads/NVIDIA_Data.xlsx")
# Load S&P500 data
sp500_data <- read_excel("~/Downloads/S&P500_Data.xlsx")

# Ensure date formatting and combine data into a time-series compatible format
nvidia_data$Date <- as.Date(nvidia_data$Date)
sp500_data$Date <- as.Date(sp500_data$Date)

# Merge datasets on date (if needed)
data_combined <- merge(nvidia_data, sp500_data, by = "Date")
```

## Plotting Time Series and Analyzing ACF/PACF

Here, I plot the time series for NVIDIA and S&P 500 closing prices to visually inspect any apparent relationship between the two. I also calculate and plot the ACF and PACF to analyze potential autocorrelations in each series.

```{r}
# Plot time-series
ggplot(data_combined, aes(x = Date)) +
  geom_line(aes(y = NVIDIA_Close, color = "NVIDIA")) +
  geom_line(aes(y = SP500_Close, color = "S&P 500")) +
  labs(title = "Time Series of NVIDIA and S&P 500", y = "Closing Price") +
  theme_minimal()

# ACF and PACF
par(mfrow = c(2, 2))
acf(data_combined$NVIDIA_Close, main = "NVIDIA ACF")
pacf(data_combined$NVIDIA_Close, main = "NVIDIA PACF")
acf(data_combined$SP500_Close, main = "S&P 500 ACF")
pacf(data_combined$SP500_Close, main = "S&P 500 PACF")
```

## STL Decomposition

I perform STL decomposition on both time series to break down each into trend, seasonal, and remainder components. This helps identify underlying patterns in the data.

```{r}
# Convert to time series and perform STL decomposition
nvidia_ts <- ts(data_combined$NVIDIA_Close, frequency = 12)
sp500_ts <- ts(data_combined$SP500_Close, frequency = 12)

nvidia_stl <- stl(nvidia_ts, s.window = "periodic")
sp500_stl <- stl(sp500_ts, s.window = "periodic")

plot(nvidia_stl, main = "NVIDIA STL Decomposition")
plot(sp500_stl, main = "S&P 500 STL Decomposition")
```

## Fitting ARIMA Models

Using auto.arima, I fit ARIMA models for NVIDIA and S&P 500 to capture trend, seasonality, and cycles, adjusting for differencing and periodic components.

```{r}
# Fit a model (e.g., ARIMA with trend and seasonality)
nvidia_arima <- auto.arima(nvidia_ts, seasonal = TRUE)
sp500_arima <- auto.arima(sp500_ts, seasonal = TRUE)

summary(nvidia_arima)
```

Here we can see that the ARIMA of the NVIDA stock is displayed as ARIMA(5,2,0)(0,0,2). This ARIMA(5,2,0)(0,0,2)[12] model uses five autoregressive terms and two levels of differencing to capture non-seasonal patterns. The model includes two seasonal moving average terms with a yearly seasonality period (for monthly data), which helps in capturing any repeating annual pattern without additional differencing for the seasonal component.

```{r}
summary(sp500_arima)
```

The ARIMA(2,1,2)(0,0,2)[12] model uses two autoregressive and two moving average terms with one differencing step to handle non-seasonal patterns. The seasonal component has two moving average terms with a 12-month period, capturing yearly seasonality without extra seasonal differencing. This setup effectively models short-term dependencies and annual patterns in the data.

## Residuals Analysis

I plot the residuals of each ARIMA model to evaluate the model’s fit. Analyzing residual patterns helps assess if further adjustments are needed.

```{r}
# Residuals vs Fitted
par(mfrow = c(1, 2))
plot(nvidia_arima$residuals, main = "NVIDIA Residuals")
plot(sp500_arima$residuals, main = "S&P 500 Residuals")
```

## 12-Step Forecast

Using the fitted ARIMA models, I forecast 12 steps ahead and plot the forecasts with error bands to visualize projected values and uncertainties.

```{r}
nvidia_forecast <- forecast(nvidia_arima, h = 12)
sp500_forecast <- forecast(sp500_arima, h = 12)

# Plot forecasts with error bands
plot(nvidia_forecast, main = "NVIDIA 12-Step Forecast")
plot(sp500_forecast, main = "S&P 500 12-Step Forecast")
```

## Fitting a VAR Model

I prepare the data for a VAR (Vector Autoregressive) model and fit it with lag order 2. This model allows us to analyze the relationship between NVIDIA and S&P 500 over time.

```{r, fig.width = 8, fig.height = 8}
# Prepare data for VAR
combined_ts <- cbind(NVIDIA = nvidia_ts, SP500 = sp500_ts)
var_model <- VAR(combined_ts, p = 2)

summary(var_model)
plot(var_model)
```

## Impulse Response Function (IRF)

To understand the impact of shocks to NVIDIA on S&P 500 and vice versa, I compute and plot the impulse response functions (IRF) of the VAR model.

```{r}
irf_results <- irf(var_model)
plot(irf_results)
```

## Granger Causality Test

I perform a Granger causality test to determine if changes in NVIDIA's series can predict changes in the S&P 500 series.

```{r}
granger_test <- causality(var_model, cause = "NVIDIA")
print(granger_test)
```

## CUSUM Test for Stability

Finally, I conduct the CUSUM test on the residuals of each series in the VAR model to check for structural stability.

```{r, fig.width=10, fig.height=6}
# Load necessary package
library(strucchange)

# Extract residuals for NVIDIA and S&P 500 from the VAR model
nvidia_resid <- residuals(var_model)[, "NVIDIA"]
sp500_resid <- residuals(var_model)[, "SP500"]

# Apply the CUSUM test on the residuals of each series
nvidia_cusum <- efp(nvidia_resid ~ 1, type = "OLS-CUSUM")
sp500_cusum <- efp(sp500_resid ~ 1, type = "OLS-CUSUM")

# Plot the CUSUM results for both residuals
par(mfrow = c(1, 2))
plot(nvidia_cusum, main = "CUSUM Test for NVIDIA Residuals")
plot(sp500_cusum, main = "CUSUM Test for S&P 500 Residuals")
```