---
title: "ECON 144 Proj 3"
author: "Leonard Zhu"
date: "2024-12-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)
library(tidyverse)
library(prophet)
library(tsibble)
library(fable)
```

# I. Introduction

The S&P 500 Index is a widely recognized benchmark for the U.S. stock market, representing the performance of 500 large-cap companies. This project analyzes daily S&P 500 closing values for forecasting, focusing on trends, seasonality, and predictive accuracy. Data was retrieved from Yahoo Finance using the quantmod package in R.

```{r}
library(quantmod)

# Fetch S&P 500 data
getSymbols("^GSPC", src = "yahoo", from = "2010-01-01", to = "2023-12-31")
sp500 <- Cl(GSPC)  # Closing prices
sp500_ts <- ts(sp500, frequency = 252)  # Convert to time series (252 trading days/year)

# Plot the data
autoplot(sp500_ts) +
  labs(title = "S&P 500 Closing Prices (2010-2023)", x = "Year", y = "Price")
```

# II. Results

## 1. ARIMA Model

```{r}
# Fit ARIMA model
arima_fit <- auto.arima(sp500_ts)
summary(arima_fit)

# Forecast with ARIMA
arima_forecast <- forecast(arima_fit, h = 252)  # Forecast 1 year ahead
autoplot(arima_forecast) +
  labs(title = "ARIMA Forecast", x = "Year", y = "Price")
```

## 2. ETS Model

```{r}
# Fit ETS model
ets_fit <- ets(sp500_ts)
summary(ets_fit)

# Forecast with ETS
ets_forecast <- forecast(ets_fit, h = 252)
autoplot(ets_forecast) +
  labs(title = "ETS Forecast", x = "Year", y = "Price")
```

## 3. Holt-Winters Model

```{r}
# Fit Holt-Winters model
hw_fit <- HoltWinters(sp500_ts)
hw_forecast <- forecast(hw_fit, h = 252)
autoplot(hw_forecast) +
  labs(title = "Holt-Winters Forecast", x = "Year", y = "Price")
```

## 4. NNETAR Model

```{r}
# Fit Neural Network model
nnetar_fit <- nnetar(sp500_ts)
summary(nnetar_fit)

# Forecast with NNETAR
nnetar_forecast <- forecast(nnetar_fit, h = 252)
autoplot(nnetar_forecast) +
  labs(title = "NNETAR Forecast", x = "Year", y = "Price")
```

## 5. Prophet Model

```{r}
# Prepare data for Prophet
sp500_df <- data.frame(ds = index(sp500), y = as.numeric(sp500))
prophet_fit <- prophet(sp500_df)

# Future data frame
future <- make_future_dataframe(prophet_fit, periods = 252)
forecast_prophet <- predict(prophet_fit, future)

# Plot Prophet Forecast
prophet_plot <- plot(prophet_fit, forecast_prophet) +
  labs(title = "Prophet Forecast", x = "Year", y = "Price")
prophet_plot
```

## 6. Forecast Combination

```{r}
# Combine forecasts using average
combined_forecast <- (arima_forecast$mean + ets_forecast$mean + hw_forecast$mean) / 3
autoplot(sp500_ts) +
  autolayer(combined_forecast, series = "Combined Forecast") +
  labs(title = "Forecast Combination", x = "Year", y = "Price")
```

## 7. Model Performance Evaluation

```{r}
# Split data into training and testing sets
# Determine the actual data range
end_year <- floor(end(sp500_ts)[1])
train <- window(sp500_ts, end = c(end_year - 1, 252))  # Train up to last complete year
test <- window(sp500_ts, start = c(end_year, 1))       # Test on the last year


# Calculate errors
arima_errors <- accuracy(forecast(auto.arima(train), h = length(test)), test)
ets_errors <- accuracy(forecast(ets(train), h = length(test)), test)
nnetar_errors <- accuracy(forecast(nnetar(train), h = length(test)), test)

# Combine results
error_df <- data.frame(
  Model = c("ARIMA", "ETS", "NNETAR"),
  RMSE = c(arima_errors[2], ets_errors[2], nnetar_errors[2])
)
knitr::kable(error_df, caption = "Model Performance Metrics")
```

# III. Conclusions and Future Work

The ARIMA model provided the lowest RMSE, making it the preferred model for this analysis. Future work could include incorporating exogenous variables (e.g., interest rates) to improve predictive accuracy.

